Comprehensive Quality Awareness Automated Semantic Web Service Composition

### 2017-03-29 paper.tex
up

In practice, many different DAGs can meet the given requirements but differ significantly in terms of QoS and semantic matchmaking quality. For example, in the classical travel planning context, during the process of building a DAG, a component service needs to be employed to obtain a travel map. Suppose that two optional services can be considered for this purpose. One service S1 can provide a street map at a price of 6.72. The other service S2 can provide a tourist map at a price of 16.87. Because in our DAG a tourist map is more desirable than a street map, service S2 clearly enjoys better semantic matchmaking quality but will have negative impact on the QoS of the DAG (i.e. the price is much higher). It can be easily imagined that similar problems can frequently occur while building the DAG for the travel planning problem or other web service composition problems. Due to this reason, aimed at achieving a good balance in between QoS and semantic matchmaking quality, our PSO-based service composition system will be further guided by a comprehensive quality model to be proposed in Section 4.

@@

\subsection{Objective Function}
In real life, given a unique and optimised solution is always easier for customers to pick up directly when many quality criteria involved into decision making, rather than provided a set of solutions. It is very practical to define a single fitness as a weighted sum of all the quality related components in Formula (\ref{equation8}). Note that weights can be adjusted according to users' preferences. The function value of 1 means the best comprehensive quality and 0 means the worst. For this purpose, $MT$, $S$, $A$, $R$, $T$, and $C$ must be normalised so that the function value falls within the range from 0 to 1 using Formula (\ref{equation9}) and (\ref{equation10}), where the maximum and minimum value of $A$, $R$, $T$, and $C$ are calculated by all web services candidates to the composition task, which are generated by a greedy search discussed in Sect.\ref{PSO_based_approach}. $MT$ and $S$ are using a bound from 0 to 1. Therefore, the composition task is try to find maximised value of objective function associated to the solutions.

@@

PSO has shown its promising in solving combinatorial optimisation problems, which is considered to an easier way to maintain the correctness of solutions comparing to GP-based approaches that often require repairing or penalising the solutions \cite{da2016particle}. Therefore, we will employ a PSO-based approach to comprehensive quality-aware automated semantic web service composition consisting a forward decoding process. Fig. \ref{overview} shows an overview of our approach consisting of four steps:

Step 1: The composition process is triggered by a composition task, which is clearly defined in Subsection \ref{problemDes}.

Step 2: This composition task is used to discover all task-related service candidates using a greedy search adopted from \cite{ma2015hybrid}. During this greedy search, an ontology-based cache ($Cache$) is initialised, which stores the concept similarities of inputs and outputs of task-related candidates.

Step 3 and Step 4: These two steps follow the stardard PSO steps \cite{shi2001particle} except some differences that lie in  particles initialisation and decoding process. These two differences are related to sort a service queue using a serivce-to-index mapping for a particle' position vectors and evaluate the fitness of a particle after decoding this serivce queue into a weighted DAG respectively, which are further addressed in Algorithm \ref{novelSteps} and \ref{graph_building} in SubSect \ref{POS-based_algomargin}.

\subsection{PSO-based approach algorithm}\label{POS-based_algomargin}
The overall algorithm investigated here is made up of the PSO-based web service composition algorithm \ref{novelSteps} and the decoding algorithm \ref{graph_building}. In algorithm \ref{novelSteps}, the  steps $4$, $5$, $6$ and $7$ are different steps from those of stardard PSO. First, each particle's position size is determined by the size of task-related service candidates generated by a greedy search. This greedy search algorithm keeps adding outputs of the evocable services as available outputs, and these available outputs are used to discovery task-related services from service repository and be updated with the outputs of these discovered services. This loop process does not stop until no service is satisfied by the available outputs. Second, each service candidate is mapped to an index of a particle’s position vector, where each vector have a weight value that falls between 0.0 and 1.0. This a service-to-index mapping for position vectors discussed in Step 3 in SubSect \ref{PSO_based_approach}. Third, all the service candidates are put in a queue and sorted according to their corresponding weight values in a descending order. Fourth, this sorted queue is used as one of the input of a forward decoding algorithm \ref{graph_building} to create a weighted DAG, this fitness value of this weighted DAG is the fitness value of the particle calculated by the comprehensive model discussed in Sect \ref{Comprehensive_Quality_Model}.
\begin{algorithm}
 %\LinesNumbered
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \SetNlSty{}{}{:}
 Randomly initialise each particle in the swarm\;
  \While {max. iterations not met}{
     \ForEach{particles in the swarm}{
     Map service candidates to particle's position vector\;
     Create a service queue and sort it by position vectors' weights\;
     Create the service into a weighted DAG \;
	 Calculate the $Weighted DAG$ fitness value\;

@@

This Algorithm  \ref{graph_building} is a forward graph graph buidling algorithm based on \cite{blum1997fast}. This algorithm take one input as the sorted service queues from step 5 of Algorithm \ref{novelSteps}. If service queues are sorted with different service order, it is possible to create different corresponding weighted DAGs as composition solutions. In addition. $I_{T}$, $O_{T}$ and $Cache$ are also taken as inputs. Firstly, $Start$ and $End$ are added to V of WG as an initialisation, and a $OutputSet$ is also created with $I_{T}$.  If all the inputs $I_{S}$ of first popped  $S$ from $queue$ can be satisfied by provided outputs of $OutputSet$, this $S$ is added to $V$ and its outputs are added to $OutputSet$. Meanwhile, $e$ is created with its $sm_{e}$ calculated based on the aggregation value of all semantic matchmaking quality from each satisfied input of $S$. These steps are not stopped until all  $O_{T}$ can satisfied from $Outputset$ or queue is null. Consequently, this forward graph building technique could lead to more services and edges connected to the $WG$, which should be removed before $WG$ is returned.

@@

\section{Experiment Study}\label{experiment_design}
In this section, a quantitative evaluation approach is adopted in our experiment design. The objectives of the evaluation are to $(1)$ measure the effectiveness of the comprehensive quality model in automated semantic web service composition approach; and $(2)$ compare solutions generated by a recent QoS-ware evaluation approach \cite{ma2015hybrid,da2016genetic} with our evaluation method proposed in the paper; and $(3)$ compare our PSP-based method with one existing GP-based approach \cite{ma2015hybrid}.

Web service challenge 2009 (WSC09) \cite{kona2009wsc} is a benchmark dataset consisting of five composition tasks that associated with an increasing number of services, and concepts in ontologies. An augmented version of WSC09 is utilised including QoS attributes for services, this version of WSC09 dataset is used in \cite{ma2015hybrid,da2016genetic}

The parameters were chosen based on general settings from \cite{shi2001particle} for our PSO-based approach, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure weight of fitness function to properly balance functional side and nonfunctional side. Therefore, $w_{1}$ and $w_{2}$ are set equally to 0.25, and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125 accordingly. The $type(a,b)$ is set to 1 ($Exact$) and 0.75 ($Plugin$) according to \cite{lecue2009optimizing}. In general, weight settings and parameter match type quality are decided by users' preferences.

\subsection{Comparison Test for GP-based approach and PSO-based approach}\label{comparisonTestWithGP}
To evaluate the performance of our proposed PSO-based approach, we compare one GP-based approach  \cite{ma2015hybrid} with our PSO-based method. we consider the links between child nodes and parent nodes as semantic web service links for measuring  semantic matchmaking quality. Therefore, we evaluate both semantic matchmaking quality and QoS simultaneously using our proposed comprehensive quality model. To make a fair comparison with our PSO-based, we consider the same number of evaluations (3000 times) used in our PSO-based approach, we set the parameters settings for the GP-based approach using 30 individuals and 100 generations.

The first column of the table \ref{meanFitness} shows five tasks from WSC09 Dataset. The second and third column of the table \ref{meanFitness} show the orginal service repository size and shrunk service repository size after the greedy search respectively regarding the five tasks for the initialisation discussed in step 2 SubSect.\ref{PSO_based_approach}. These significant decreased number of services demonstrate that the researching space is effectively reduced. The fourth and fifth column of the table \ref{meanFitness} show the mean fitness values of 30 independent runs accomplished by two methods. We employ statistical analysis to test the significant differences in mean fitness value. The results show that the PSO-based approach performs better in four of five tasks (all the p-values are consistently smaller than 0.01).

\begin{table}[]
\centering
\caption{Mean fitness results for comparing GP-based approach}
\label{meanFitness}
\begin{tabular}{l|l|l|l|l}
\hline
\multicolumn{1}{c|}{WSC09} &Original $SR$  &Shrunk $SR$   &PSO-based approach & GP-based approach  \\ \hline
Task 1                     &572            &80    &0.5592 $\pm$ 0.0128  $\uparrow$  &0.5207 $\pm$ 0.0208           \\ \hline
Task 2                     &4129           &140   &0.4701 $\pm$ 0.0011  $\uparrow$  &0.4597 $\pm$ 0.0029          \\ \hline
Task 3                     &8138           &153   &0.5504 $\pm$ 0.0128              &0.5679 $\pm$ 0.0234 $\uparrow$   \\ \hline
Task 4                     &8301           &330   &0.4690 $\pm$ 0.0017  $\uparrow$  &0.4317 $\pm$ 0.0097            \\ \hline
Task 5                     &15211          &237   &0.4694 $\pm$ 0.0008  $\uparrow$  &0.2452 $\pm$ 0.0369            \\ \hline
\end{tabular}
\end{table}

@@

To better understand a better semantic matchmaking quality with a slight trade-off in the opitmal solutions using comprehensive quality evaluation model, we demonstrate the optimal solutions using task 3 as an example. Fig. \ref{comparisontest} $(1)$ and $(2)$ show two weighted DAGs as best solutions found by employing QoS evaluation model and comprehensive quality evaluation model respectively. Two approaches generate exactly the same service workflow structure, but some service vertices and edges denoted in red are different. To better understand these differences, we list the overall semantic matchmaking quality $SM$,  overall $QoS$ and semantic matchmaking quality associated to each edge ($sm_{e_1}$ to $sm_{e_4}$) for the two weighted DAGs in Fig. \ref{comparisontest} $(3)$, where $\Delta Q$ reveals the gain (positive $\Delta Q$) or a loss (negative $\Delta Q$) for the listed qualities using our comprehensive quality evalutaion model. Therefore, an overall gain 0.1433 is calculated from a sum of a $SM$ gain (0.1467) and $QoS$ loss (-0.0034). Consequently, our comprehensive evaluation model acheive a desirable trade-off in considering both $SM$ and $QoS$. To explain the value of $SM$ gain, we pick up $e_4$ that is associated with a smallest $\Delta Q$. The $e_4$ of $WG(1)$ using QoS evaluation model and $WG(2)$ using comprehensive quality model has two different source service vertices $Ser1640238160$ and $Ser947554374$ respectively,  and the same $end$ vertices. $Ser1640238160$ and $Ser947554374$ are services with concept-related output parameters $Inst795998200$ and $Inst582785907$ corresponds to two concepts $Con103314376$ and $Con2037585750$ respectively, which are marked on the related taxonomy in Fig. \ref{comparisontest} $(4)$. In addtion, $Inst658772240$ is a required parameter of the $end$ vertice as one of composition task outputs, which is related to concept $Con2113572083$. There exist $Inst582785907$ $\in$ $Con2037585750$ $\sqsubseteq_{3}$ $Inst658772240$ $\in$ $Con2113572083$ for $e_4$ in $WG(1)$ and $Inst795998200$ $\in$ $Con103314376$ $\sqsubseteq_{2}$ $Inst658772240$ $\in$ $Con2113572083$ for $e_4$ in $WG(2)$. Therefore, 2-level Subsumes is closer than 3-level Subsumes, $Inst795998200$ is closer to user's required output $Inst658772240$ compared with $Inst582785907$.

>>>>

In practice, many different DAGs can meet the given requirements but differ significantly in terms of QoS and semantic matchmaking quality. For example, in the classical travel planning context, during the process of building a DAG, a component service needs to be employed to obtain a travel map. Suppose that two optional services can be considered for this purpose. One service S1 can provide a street map at a price of 6.72. The other service S2 can provide a tourist map at a price of 16.87. Because in our DAG a tourist map is more desirable than a street map, service S2 clearly enjoys better semantic matchmaking quality but will have negative impact on the QoS of the DAG (i.e. the price is much higher). It can be easily imagined that similar problems can frequently occur while building the DAG for the travel planning problem or other web service composition problems. Due to this reason, aimed at achieving a good balance in between QoS and semantic matchmaking quality, our PSO-based service composition system will be further guided by a comprehensive quality model to be proposed in 4.

@@

\subsection{Objective Function}
In real life, given a unique and optimised solution is always easier for customers to pick up directly when many quality criteria involved into decision making, rather than provided a set of solutions. It is very practical to define a single fitness as a weighted sum of all the quality related components in Formula (\ref{equation8}). Note that weights can be adjusted according to users' preferences. The function value of 1 means the best comprehensive quality and 0 means the worst. For this purpose, $MT$, $S$, $A$, $R$, $T$, and $C$ must be normalised so that the function value falls within the range from 0 to 1 using Formula (\ref{equation9}) and (\ref{equation10}), where the maximum and minimum value of $A$, $R$, $T$, and $C$ are calculated by all web services candidates to the composition task, which are generated by a greedy search discussed in \ref{PSO_based_approach}. $MT$ and $S$ are using a bound from 0 to 1. Therefore, the composition task is try to find maximised value of objective function associated to the solutions.

@@

PSO has shown promise in solving combinatorial optimisation problems, and considered an easier way to maintain the correctness of solutions compared to GP-based approaches that often require repairing the solutions \cite{da2016particle}. Therefore, we employ a PSO-based approach to comprehensive quality-aware automated semantic web service composition. Fig. \ref{overview} shows an overview of our approach consisting of four steps:

Step 1: The composition process is triggered by a composition task, which is clearly defined in \ref{problemDes}.

Step 2: This composition task is used to discover all task-related service candidates using a greedy search adopted from \cite{ma2015hybrid}. This greedy search algorithm keeps adding outputs of the invoked services as available outputs (initialised with $I_{T}$) , and these available outputs are used to discover task-related services from a service repository and updated with the outputs of these discovered services. This operation is repeated until no service is satisfied by the available outputs. This greedy search contributes to a shrunken service repository. During the greedy search, an ontology-based cache ($cache$) is initialised that stores the concept similarities of potentially matched inputs and outputs of task-related candidates. This $cache$ is also used to check whether services can be invoked by satisfying their input-related concepts with provided output-related concepts.

Step 3 and Step 4: These two steps follow the standard PSO steps \cite{shi2001particle} except for some differences in particles mapping and decoding processes. In particular, these two differences are related to sorting a created service queue using serivce-to-index mapping for a particle' position vectors and evaluating the fitness of a particle after decoding this serivce queue into a weighted DAG respectively. Those differences are further addressed in Algorithms \ref{novelSteps} and \ref{graph_building} in \ref{POS-based_algomargin}.

\subsection{PSO-based approach algorithm}\label{POS-based_algomargin}
The overall algorithm investigated here is made up of a PSO-based web service composition algorithm \ref{novelSteps} and a decoding algorithm \ref{graph_building}. In Algorithm \ref{novelSteps}, the  steps $4$, $5$, $6$ and $7$ are different from those of stardard PSO: In step 4, the size of task-related service candidates generated by a greedy search determines the size of each particle's position, and each candidate in a created service candidates queue is mapped to an index of a particle’s position vectors, where each vector has a weight value between 0.0 and 1.0. In step 5, all the task-related service candidates in the queue are sorted according to their corresponding weight values in descending order. In step 6, this sorted queue is used as one of the inputs of the forward decoding algorithm \ref{graph_building} to create a weighted DAG. In step 7, this fitness value of this weighted DAG is the fitness value of the particle calculated by the comprehensive model discussed in \ref{Comprehensive_Quality_Model}.
\begin{algorithm}
 %\LinesNumbered
 \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
 \SetKwFunction{generateWeightedGraph}{generateWeightedGraph}
 \SetKwProg{Procedure}{Procedure}{}{}
 \SetNlSty{}{}{:}
 Randomly initialise each particle in the swarm\;
  \While {max. iterations not met}{
     \ForEach{particles in the swarm}{
     Create a service candidates queue and map service candidates to a particle's position vectors\;
     Sort the service queue by position vectors' weights\;
     Create a weighted DAG from the service queue ( Algorithm\ref{graph_building}) \;
     Calculate the Weighted DAG fitness value\;

@@

Algorithm  \ref{graph_building} is a forward graph buidling algorithm based on \cite{blum1997fast}. This algorithm take one input, a sorted service queue from step 5 of Algorithm \ref{novelSteps}. If service queues are sorted resulting in different service order, it is possible to create different corresponding weighted DAGs as composition solutions. In addition. $I_{T}$, $O_{T}$ and $cache$ are also taken as the inputs. Firstly, $Start$ and $End$ are added to $V$ of $WG$ as an initialisation, and $OutputSet$ is also created with $I_{T}$.  If all the inputs $I_{S}$ of the first popped  $S$ from $queue$ can be satisfied by provided outputs from $OutputSet$. This $S$ is added to $V$ and its outputs are added to $OutputSet$, and $S$ is removed from $queue$. Meanwhile, $e$ is created with its $sm_{e}$ calculated based on the aggregation value of all semantic matchmaking qualities from each satisfied input of $S$. These steps are repeated until all  $O_{T}$ can be satisfied by $Outputset$ or the service queue is $null$. Consequently, this forward graph building technique could lead to more services and edges connected to the $WG$, which should be removed before $WG$ is returned.

@@

\section{Experiment Study}\label{experiment_design}
In this section, a quantitative evaluation approach is adopted in our experiment design. Two objectives of this evaluation are to: $(1)$ measure the effectiveness of our PSO-based approach. To do this, we compare our PSO-based method with one existing GP-based approach \cite{ma2015hybrid}. $(2)$ measure the effectiveness of our proposed comprehensive quality model. To do this, we compare a widely used QoS model in QoS-aware web service composition \cite{ma2015hybrid,da2016particle,da2015graphevol} with our comprehensive quality model using our proposed PSO-based approach.

An augmented version of Web service challenge 2009 (WSC09) used in \cite{ma2015hybrid,da2016genetic} including QoS attributes is used for our experiment. WSC09 is a benchmark dataset consisting of five composition tasks that associated with an increasing number of services, and concepts in ontologies.

The parameters are chosen based on the settings from \cite{shi2001particle} for our PSO-based approach, In particular, PSO population size is 30 with 100 generations. We run 30 times independently for each dataset. We configure weight of fitness function to properly balance functional side and nonfunctional side. Therefore, $w_{1}$ and $w_{2}$ are set equally to 0.25, and $w_{3}$, $w_{4}$, $w_{5}$, $w_{6}$ are all set to 0.125. The returned value of $type(a,b)$ is set to 1 ($Exact$) and 0.75 ($Plugin$) according to \cite{lecue2009optimizing}. In general, weight settings and parameter match type quality are decided by users' preferences.

\subsection{Comparison Test for GP-based approach and PSO-based approach}\label{comparisonTestWithGP}
To evaluate the effectiveness of our proposed PSO-based approach, we compare one recent GP-based approach \cite{ma2015hybrid} with our PSO-based method. The semantic matchmaking quality of this GP-based approach is easy to evaluated considering measuring links between parent nodes and children nodes. Therefore, we evaluate both semantic matchmaking quality and QoS simultaneously for that GP-based approach using the proposed comprehensive quality model. To make a fair comparison, we consider the same number of evaluations (3000 times) used in our PSO-based approach. We set the parameters' settings of that GP-based approach as 30 individuals and 100 generations, it is considered to be proper settings refering to \cite{da2015gp}.

The first column of Table \ref{meanFitness} shows five tasks from WSC09 Dataset. The second and third column of Table \ref{meanFitness} show the orginal service repository size before the greedy search and shrunk service repository size after the greedy search respectively regarding the five tasks. This greedy search helps reducing the repository size by considering only task-related service candidates, which also contributes to a significant reduced researching space. The fourth and fifth column of Table \ref{meanFitness} show the mean fitness values of 30 independent runs accomplished by two methods. We employ independent-samples T tests to test the significant differences in mean fitness value. The results show that the PSO-based approach outperformes the existing GP-based approach in most cases except task 3 (all the p-values are consistently smaller than 0.01). In task 5, PSO-based approach performs significantly better than GP-based approach in finding optimal solutions. It may refer to that GP-based approach is stuck in the local optimal due to a very large searching space of our composition problem, but PSO-based approach requres a decoding process, and small changes in a sorted queue might lead to a varying decoded solutions. Therefore, PSO-based approach is considered to be hard to be stuck in the local optimal for our composition problem.

\begin{table}[]
\centering
\caption{Mean fitness results for comparing GP-based approach}
\label{meanFitness}
\begin{tabular}{l|l|l|l|l}
\hline
\multicolumn{1}{c|}{WSC09} &Original $SR$  &Shrunken $SR$   &PSO-based approach & GP-based approach  \\ \hline
Task 1                     &572            &80    &0.5592 $\pm$ 0.0128  $\uparrow$  &0.5207 $\pm$ 0.0208           \\ \hline
Task 2                     &4129           &140   &0.4701 $\pm$ 0.0011  $\uparrow$  &0.4597 $\pm$ 0.0029          \\ \hline
Task 3                     &8138           &153   &0.5504 $\pm$ 0.0128              &0.5679 $\pm$ 0.0234 $\uparrow$   \\ \hline
Task 4                     &8301           &330   &0.4690 $\pm$ 0.0017  $\uparrow$  &0.4317 $\pm$ 0.0097            \\ \hline
Task 5                     &15211          &237   &0.4694 $\pm$ 0.0008  $\uparrow$  &0.2452 $\pm$ 0.0369            \\ \hline
\end{tabular}
\end{table}

@@

To better understand a better semantic matchmaking quality with a slight trade-off in the opitmal solutions using comprehensive quality evaluation model, we demonstrate the optimal solutions using task 3 as an example. Fig. \ref{comparisontest} $(1)$ and $(2)$ show two weighted DAGs as best solutions found by employing QoS evaluation model and comprehensive quality evaluation model respectively. Two approaches generate exactly the same service workflow structure, but some service vertices and edges denoted in red are different. To better understand these differences, we list the overall semantic matchmaking quality $SM$,  overall $QoS$ and semantic matchmaking quality associated to each edge ($sm_{e_1}$ to $sm_{e_4}$) for the two weighted DAGs in Fig. \ref{comparisontest} $(3)$, where $\Delta Q$ reveals the gain (positive $\Delta Q$) or a loss (negative $\Delta Q$) for the listed qualities using our comprehensive quality evalutaion model. Therefore, an overall gain 0.1433 is calculated from a sum of a $SM$ gain (0.1467) and $QoS$ loss (-0.0034). Consequently, our comprehensive evaluation model acheive a desirable trade-off in considering both $SM$ and $QoS$. To explain the value of $SM$ gain, we pick up $e_4$ that is associated with a smallest $\Delta Q$. The $e_4$ of $WG(1)$ using QoS evaluation model and $WG(2)$ using comprehensive quality model has two different source service vertices $Ser1640238160$ and $Ser947554374$ respectively,  and the same $end$ vertices. $Ser1640238160$ and $Ser947554374$ are services with concept-related output parameters $Inst795998200$ and $Inst582785907$ corresponds to two concepts $Con103314376$ and $Con2037585750$ respectively, which are marked on the related taxonomy in Fig. \ref{comparisontest} $(4)$. In addtion, $Inst658772240$ is a required parameter of the $end$ vertice as one of composition task outputs, which is related to concept $Con2113572083$. Obviouly,  $Inst795998200$ is closer to user's required output $Inst658772240$ compared to $Inst582785907$.
